### 2022-03-31 Notes/overview.tex

\section{overview, general flow...}

\subsection{GBW\cite{gbw1998}}
for flavour $f$ and polarization $p$ \cite{gbw1998} \cite{cambqcd},
\begin{equation}
\sigma_{f,\;p}=\int_{0}^{1} d z \int d^2\mathbf{r}|\Psi_{f,\;p}|^2 \sigma_{dp}.
\end{equation}

photon wavefunctions are known (cf.  \cite{gbw1998}).

Original GBW model introduces $\sigma_{dp}$ in the form
\begin{equation}
\sigma_{GBW} (x, r )=\sigma_0 (1-e^{-\hat{r}^2}),
\end{equation}
where $\hat{r}^2$ which is the main object here to be defined as $\frac{r^2}{4 R^2_0(x)}$, with saturation scale \hyperlink{saturationscaleGBW}{$R^2_0$}, and modified \hyperlink{modifiedx}{$x$}.
In this model, there are three parameters to be fitted, namely $\sigma_0$, $\lambda$ and $x_0$.

\subsection{GBW-BK \cite{gbk2002}}
GBW-BK model introduces diple cross section $\sigma_{dp}$ in the form
\begin{equation}
\sigma_{GBWBK}(x,r)=\sigma_0\left(1-\exp\left(-\frac{\pi^2 r^2\alpha_s(\mu^2) x g(x,\mu^2)}{3\sigma_0}\right) \right)
\end{equation}

One important feature is,
\begin{equation}
\lim_{r \rightarrow 0}\sigma_{GBWBK} \simeq  \frac{\pi^2 r^2\alpha_s(\mu^2) x g(x,\mu^2)}{3},
\end{equation}
where $g(x,\mu^2)$ is integrated gluon density. This is the limit $r \rightarrow 0$ of \hyperlink{ugd}{ $\sigma_{dp}$ in terms of unintegrated gluon density}.

The scale is proposed to be $\mu^2=\frac{C}{r^2}+{\mu^2_0}$.

\subsection{GBW-BK-S}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{notation, miscelaneous}

$k_g$ transverse momentum of gluon.\\
$k_q$ transverse momentum of quark.\\
$r$  dipole size.\\
dipole scattering amplitude dissapears
 for $r=0$ (cf. color trnsparency), and saturates to a constant for large $r$. cf.\cite{gbw1998}.

\begin{equation}
\hypertarget{modifiedx}{x}=x_{Bjorken}\left(1+\frac{4m^2_q}{Q^2}\right),
\end{equation}
where $m_q$ is effective quark mass.

The saturation scale in GBW \hypertarget{saturationscaleGBW}{$R^2_0$}, becomes in the limit $x\rightarrow 0$
\begin{equation}
R_0\rightarrow \frac{1}{GeV^2}\left(\frac{x}{x_0}\right)^\lambda,
\end{equation}
where $x_0$ and $\lambda$ to be fitted.

Scale $\mu^2$ in GBW-BK behaves like $\frac{const}{r^2}$ in small r region, and like a constant in large $r$ region.

Unintegrated gluon density\hypertarget{ugd}{$f(x,k^2)$} is related to the dipole cross section $\sigma_{dp}$ by% the fourier trransform
\cite{gbk2002, nikolaev1994, nikolaev2001}:

\begin{equation}
\sigma_{dp}=\frac{2 \pi}{3} \int\frac{d^2 \mathbf{k_g}}{k_g^4} \alpha_s f(x,k_g^2) \left(1-e^{i \pairdot{k_g}{r}}\right).
\end{equation}

and hence the fourier transform
\begin{align}
\nabla^2_{\mathbf{r}} \sigma_{dp}&=\frac{2 \pi}{3} \int\frac{d^2 \mathbf{k_g}}{k_g^2} \alpha_s f(x,k_g^2)  e^{i \pairdot{k_g}{r}}.
\end{align}

In the small $r$ limit\cite{nikolaev2001}, 
\begin{equation}
\sigma_{dp}(x,r)=\frac{\pi^2 r^2\alpha_s(\mu^2=\frac{C}{r^2}) x g(x,\mu^2=\frac{C}{r^2})}{3}.
\end{equation}

unintegrated gluon density $f(x,k_g^2)$ is related to the integrated gluon density $g(x,Q^2)$ by 
 \begin{equation}
x g (x,Q^2)=\int d k_g^2 f(x,k_g^2) \theta(Q^2-k_g^2)/k_g^2.
\end{equation}