### 2023-05-08 Paper/draft.tex
documet update

\documentclass[11pt]{article}
\usepackage[a4paper,left=0.99in, right=0.99in,top=1.2in, bottom=1.2in]{geometry}

\usepackage[english]{babel}
\usepackage[a4paper]{geometry}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{breqn}
%\usepackage[]{fancyhdr}
\usepackage{tikz}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{cite}
%\usepackage{ulem}

\usetikzlibrary{decorations}
\usetikzlibrary{snakes}

%-----------------------------------------------------------------------------
%\pagenumbering{arabic}
%\numberwithin{equation}{section}

%-----------------------------------------------------------------------------
\newcommand{\fww}[0]{\mathcal{F}^{\mathrm{WW}}}
\newcommand{\fdp}[0]{\mathcal{F}^{\mathrm{dipole}}}
\newcommand{\sdp}[0]{\sigma_{\mathrm{dipole}}}
\newcommand{\sdpa}[0]{\sigma_{\mathrm{dipole\,Adj.}}}
\newcommand{\GeV}[0]{\mathrm{GeV}}
\newcommand{\comment}[1]{\texttt{\color{red}#1}}
\newcommand{\commentPending}[1]{\texttt{\color{red!25}#1}}
\newcommand{\pairdot}[2]{ \mathbf{#1}\cdot\mathbf{#2}  }

%-----------------------------------------------------------------------------
\author{
T.~Goda,$^1$, K.~Kutak$^{1}$ and S.~Sapeta$^1$\\\,\\
$^1$ 
{\small\it The H.\ Niewodnicza\'nski Institute of Nuclear Physics PAN,}\\ 
{\small\it Radzikowskiego 152, 31-342 Krak\'ow, Poland}\\
}

\title{Effects of gluon kinematics and the Sudakov form factor\\ on the dipole amplitude}

\date{}

%-----------------------------------------------------------------------------
\begin{document}

\maketitle

%--------------- preprint numbers ---------------------
\vspace{-22em}
\begin{flushright}
  IFJPAN-IV-2023-XX
\end{flushright}
\vspace{17.5em}
%------------------------------------------------------

\begin{abstract}
We investigate effects of exact gluon kinematics on the parameters of the
Golec-Biernat--W\"usthoff, and Bartels--Golec-Biernat--Kowalski saturation
models. The resulting parameters show some differences, particularly, the
normalization of the dipole cross section $\sigma_0$. The refitted models are
used for the dijet production process in DIS to investigate effects of the
Sudakov form factor at Electron Ion Collider energies.  
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

\comment{preprint number (SS)}\\

\begin{figure}[t]
\centering
\resizebox{0.6\textwidth}{0.2\textwidth}{
\begin{tikzpicture}
  \draw[snake=coil,segment aspect=0,segment length=0.1in](-0.875in , 0.5in)--(-0.5in,0.in);
  \draw[snake=coil,segment aspect=0,segment length=0.1in](0.875in ,0.5in)--(0.5in,0.in);
  \draw(-0.5in , 0in)--(0.5in,0in)--(0.5in , -0.5in)--(-0.5in,-0.5in)--cycle ;

  \filldraw(0.02in,0)--(-0.02in,-0.02in)--(-0.02in,0.02in)--cycle;
  \filldraw(-0.02in,-0.5in)--(0.02in,-0.52in)--(0.02in, -0.48in)--cycle;

  \filldraw(-0.5in,-0.23in)--(-0.52in,-0.27in)--(-0.48in,-0.27in)--cycle;
  \filldraw(0.5in,-0.27in)--(0.52in,-0.23in)--(0.48in,-0.23in)--cycle;

  \draw(-0.5in , -0.25in)node[left,scale=0.02in]{$x$, $\kappa$};
  \draw(-0.5in , -0.75in)node[left,scale=0.02in]{$x/z$, $k$};
  \draw[ snake=coil,segment aspect=1.5,segment length=0.1in](0.5in , -1in)--(0.5in,-0.5in);
  \draw[snake=coil,segment aspect=1.5,segment length=0.1in](-0.5in , -1in)--(-0.5in,-0.5in);
  \end{tikzpicture}
  \hspace{60pt}
  \begin{tikzpicture}
  \draw[snake=coil,segment aspect=0,segment length=0.1in](-0.875in , 0.5in)--(-0.5in,0.in);
  \draw[snake=coil,segment aspect=0,segment length=0.1in](0.875in ,0.5in)--(0.5in,0.in);
  \draw(-0.5in , 0in)--(0.5in,-0.5in)--(0.5in ,0in)--(-0.5in,-0.5in)--cycle ;

  \filldraw(-0.5in,-0.23in)--(-0.52in,-0.27in)--(-0.48in,-0.27in)--cycle;
  \filldraw(0.5in,-0.23in)--(0.52in,-0.27in)--(0.48in,-0.27in)--cycle;

  \draw[ snake=coil,segment aspect=1.5,segment length=0.1in](0.5in , -1in)--(0.5in,-0.5in);
  \draw[snake=coil,segment aspect=1.5,segment length=0.1in](-0.5in , -1in)--(-0.5in,-0.5in);
\end{tikzpicture}
}
\caption{Kinematic variables of the gluon contribution to $F_2$. The momentum fraction of gluon is $x/z$ as it has to contribute via the quark box.
}
\label{fig:gluon-kinem}
\end{figure}

Factorization of scales plays central role in {\it Quantum Chromodynamics}
(QCD). In particular, within collinear factorization approach, long-distance
effects can be isolated into objects called {\it collinear parton distribution
functions}~(PDFs)~\cite{Roberts:1990ww,Collins:2011zzd}.  The collinear PDFs are not
fully perturbatively calculable but they obey perturbative evolution equations
and are process independent~\cite{Collins:2011zzd}. That is to say that one needs
to measure them in experiments, but the PDFs obtained in one experiment can be
used as input in another.

One of the main type of processes that is suited particularly well to the studies
of proton structure is the {\it deep inelastic
scattering}~(DIS)~\cite{Roberts:1990ww,Collins:2011zzd}. The data from HERA has
enabled us to answer many questions in that domain and largely improved our
picture of interior of a proton. 
It is therefore very exiting that the next generation of a DIS machines,
the election-ion collider~(EIC)~\cite{NAP25171} is making its way and will soon
allow us to uncover more details about the hadron structure.

With an increase of the of the center of mass energy at fixed value of the
photon virtuality, $Q$, one shall probe  the region of the small Bjorken scaling
variable, $x$.  
In that region, dominated by gluons\cite{Roberts:1990ww,Collins:2011zzd,
Kovchegov:2012mbw}, the appropriate theoretical formalism to use is the Color
Glass Condensate~\cite{McLerran:1993ni, Ayala:1995kg, Kovchegov:1996ty,Jalilian-Marian:1996mkd,Iancu:2000hn,Iancu:2001ad,Ferreiro:2001qy,Iancu:2002xk} or, if the density of gluons is still moderate, the
High Energy Factorization also called $k_T$-factorization~\cite{Catani:1990xk,Catani:1990eg,Collins:1991ty,Catani:1993ww,Catani:1993rn,Catani:1994sq}.

In the above frameworks, the interaction of the photon with gluons happens
through a box of quarks, as depicted in Fig.~\ref{fig:gluon-kinem}. An object,
which is central to the description of this process is called the dipole gluon
density, $\fdp(x,k_T^2)$, which is a type of {\it transverse momentum
dependent} PDFs (TMD
PDFs)~\cite{Bomhof:2006dp,Dominguez:2010xd,Dominguez:2011wm,Xiao:2017yya}. 
%
It appears in the $k_T$-factorization
formula~\cite{Catani:1990eg,Dominguez:2010xd,Dominguez:2011wm,Xiao:2017yya},
which we shall discuss in the next section, and plays a role similar to that of
the collinear parton distribution function in the collinear factorization.

Much effort has been made
to study this object, notably the work of Balitsky, Fadin, Kuraev and Lipatov
(BFKL)~\cite{Balitsky:1978ic, Kuraev:1977fs} has predicted sharp rise of cross
sections with $1/x$.  However, such rise violates unitarity and the Froissart
bound~\cite{Roberts:1990ww,Collins:2011zzd,Kovchegov:2012mbw,Barone:1993sy}. In
\cite{Gribov:1983ivg} it has been recognized that gluon recombination can tame
the growth of gluons \cite{Gribov:1983ivg}. The interplay of linear and
nonlinear terms leads to a phenomenon called {\it gluon saturation} and
corresponding evolution equations are known as
Balitsky-Kovchegov~(BK)~\cite{Balitsky:1995ub,Kovchegov:1999yj} equation or
JIMWLK\cite{Kovner:1999bj,Kovner:2000pt,Iancu:2000hn,JalilianMarian:1997dw,JalilianMarian:1997gr,JalilianMarian:1997jx} equation.

Due to its clear picture, one often uses a position space counter part of the gluon density, called {\it dipole cross section}. Despite the success of the aforementioned evolution equations, there exists a number of phenomenological models of the dipole cross section, which are popular due to to their simplicity. 
\commentPending{More about saturation models}
A notable example of a dipole saturation model is the one proposed by
Golec-Biernat and Wu\"sthoff (GBW)~\cite{Golec-Biernat:1998zce} and the one proposed by
Bartels, Golec-Biernat and Kowalski (BGK)~\cite{Bartels:2002cj}. We will use these two models in this study. There are other models, with different considerations, such as impact parameter dependence~({\it cf.}
\cite{Kowalski:2003hm,McLerran:1993ni,Forshaw:1999uf,Iancu:2003ge}).

\commentPending{Section about what we did: fits of saturation models with exact
kinematics, use of those results + Sudakov for EIC predictions.}

In the present study, we fit the GBW~\cite{Golec-Biernat:1998zce} and BGK~\cite{Bartels:2002cj} models in the $k_T$ factorization formula of $F_2$, instead of the original dipole factorization formula. As discussed below, this allows us to fit the model with exact gluon kinematics. 
Since inclusive observables do not reveal the $k_T$ dependence well, we also consider a less inclusive observable which is more sensitive to the $k_T$ dependence of the gluon distribution.
The resulting models are put to the test by applying them to the predictions of dijet DIS at EIC, including effects of the Sudakov form factor, which resums logarithms of the small transverse momentum.

The paper is organized as follows.  In the next section, we present theoretical
framework which we used and, in particular, outline differences of the
$k_T$-factorization and the dipole-factorization.  In
Sec.~\ref{sec:F2fits}, results of the new fits to HERA data in the
$k_T$-factorization formula are presented. 
In Sec.~\ref{sec:dijetsEIC}, we apply the results from the previous section to
compute distributions for dijet production in DIS at the Electron-Ion Collider
(EIC)~\cite{NAP25171} and make comparisons to our earlier study of Ref.~\cite{vanHameren:2021sqc}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The framework}
\commentPending{Missing a general factorization formula}
In general, a factorization formula has the form
\begin{equation}
    \sigma=H \otimes \phi,
\end{equation}
where $\otimes$ denotes convolution. In this form one achieves two things. Firstly, one can confine long distance effects in one of the factors (let $\phi$ be such an object), and leaves $H$ as an perturbatively calculable object. Secondly, one may factorize out a (to some extent) process independent object.  
In the present study, the dipole cross section and the gluon densities both  contain nonperturbative part and are, to an extent, process independent~\footnote{They are not universal, however they may appear in different processes. (see for example Refs.~\cite{Dominguez:2011wm,Kotko:2015ura})}.
In this section, we will outline two factorizations; the $k_T$ factorization and the dipole factorization. \\
As it was mentioned earlier, we use the GBW and BGK models.  
The models were originally formulated in the position-space version of the
$k_T$-factorization formula. %, and the object which corresponds to the gluon
%density of the $k_T$-factorization is called the dipole cross section. 
In the GBW model, the dipole cross section has the form~\cite{Golec-Biernat:1998zce}
\begin{align}
\label{eq:gbw-dipole}
\sigma_{\mathrm{GBW}}(x,r)&=\sigma_0\left(1-e^{-r^2/R^2_0}\right)&\mathrm{where}&
&R^{-2}_0&=\frac{Q_0^{2}}{4}\left(\frac{x_0}{x}\right)^\lambda.
\end{align}
%
The dipole cross section is related to the gluon density by the Fourier
transform
%
\begin{equation}
\alpha_s\fdp(x,k_T^2)=\frac{N_c}{4\pi}\int\frac{d^2\mathbf{r}}{(2\pi)^2}e^{i\mathbf{k}_t\cdot \mathbf{r} }\nabla_{\mathbf{r}}^2\sigma_{\mathrm{dipole}}(x,r).
\label{eq:dipole-gluon}
\end{equation}
%
The essence of the GBW model is encoded in the $x$-dependent saturation scale
$Q_s^2(x)=Q^2_0(x_0/x)^{-\lambda}$, which separates the saturation region and
the scaling region.  A extension of the above model was proposed by Bartels, Golec-Biernat
and Kowalski~(BGK)~\cite{Bartels:2002cj} who incorporated the DGLAP evolution
in the GBW cross section~(\ref{eq:gbw-dipole}) by modifying the exponent to 
%
\begin{equation}
R_0^{-2}=\frac{\pi^2\alpha(\mu^2)xg(x,\mu^2)}{3\sigma_0},
\end{equation}
where
\begin{align}
\mu^2&=\frac{C}{r^2}+\mu_0^2 & &\mathrm{and} & xg(x,Q^2_0)&=A_g x^{-\lambda_g}(1-x)^{5.6},
\end{align}
%
thus improving the description of data at higher $Q^2$.\\

While these models enjoyed much success in phenomenology of DIS, including the
diffractive and photo-production
processes~\cite{Golec-Biernat:1998zce,Golec-Biernat:1999qor}, it is worth
mentioning that they both use certain kinematic approximation which follows from
subtle differences in the dipole factorization and the $k_T$-factorization.  In
the following, we will investigate effects of these approximations.

Firstly, let us outline the factorization formulas.
With $q'\equiv q+x p$, one decomposes $k$ and $\kappa$, defined in
Fig.~\ref{fig:gluon-kinem}, as
%
\begin{align}
    \kappa&=\alpha p-\beta q'+\kappa_t&&\mathrm{and}& k&=a p- bq'+k_T.
\end{align}
For the contribution depicted in Fig.~\ref{fig:gluon-kinem}, the structure function, $F_2$ for transversely and longitudinally polarized photons, factorize, with a change of variable  ${\boldsymbol{\kappa}'}_t\equiv{\boldsymbol{\kappa}_t}-(1-\beta)\mathbf{k}_t$, in the form~\cite{ Kimber:2001uaa,Kwiecinski:1997ee}
\begin{multline}
	F_2(x,Q^2)=\sum_f e_f^2 \frac{Q^2}{2\pi}\int\frac{dk^2_t}{k_T^2}\int^1_0d\beta\int d{\kappa'}_t\alpha_s(\mu^2) \fdp(x/z,k_T^2)\Theta(1-x/z)\\
	\left[\left(\beta^2+(1-\beta)^2\right)\left(\frac{I_1}{2\pi}-\frac{I_2}{2\pi}\right)
	+\left(m_f^2+4Q^2\beta^2(1-\beta)^2\right)\left(\frac{I_3}{2\pi}-\frac{I_4}{2\pi}\right)\right],
	\label{eq:angle-integrated}
\end{multline}
where
\begin{equation}
	\frac{1}{z}=1+\frac{{\kappa'}^2_t+m_f^2}{\beta(1-\beta)Q^2}+\frac{k^2_t}{Q^2}\,.
	\label{eq:z}
\end{equation}

One should note, that gluons are not probed directly, thus the argument of the
gluon density is $x/z$ rather than $x$.  If one, instead, uses $ 1/z=1+4
m_f^2/Q^2$ and assumes that $\mu$ is independent of $\kappa'$ and $k_T$, the
above formula can be written in the impact parameter
space~\cite{Golec-Biernat:1998zce,Nikolaev:1990ja} ({\it i.e.} as a
dipole-factorization formula)
%
\begin{equation}
  F_2\left(x,Q^2\right)=\sum_f e_f^2 \frac{Q^2}{4\pi^2} \int^1_0d\beta \int \frac{d^2\mathbf{r}}{(2\pi)^2} \left|\Psi\left(\tilde{x},\beta,Q^2\right)\right|^2\sigma_{\mathrm{dipole}}\left(\tilde{x},r\right),
  \label{eq:dipole-factorization}
\end{equation}

where the photon wave function, $\left|\Psi\left(x,\beta,Q^2\right)\right|^2$,
describes splitting of the incoming photon into a $q\overline{q}$ pair with
light-cone momenta fractions $\beta$ and $1-\beta$ respectively, and the dipole
cross section, $\sigma_{\mathrm{dipole}}\left(\tilde x,r\right)$, describes the
interaction of the colour dipole of the size $r$ with the proton. In
Ref.~\cite{Golec-Biernat:1998zce}, $\tilde{x}=x(1+4m_f/Q^2)$ with nonzero value
of $m_f$ was used even for light flavours, which is necessary to partially
substitute $k_T$ and ${\kappa'}_T$ in Eq.~(\ref{eq:z}).
%The original GBW and BGK models were fitted in the dipole factorization, Eq.~(\ref{eq:dipole-factorization}), while 
In the present study, we use Eq.~(\ref{eq:angle-integrated}) to fit the models, where the gluon density is obtained by evaluating Eq.~(\ref{eq:dipole-gluon}).
%Comparing, the GBW and BGK models, one identifies that, the GBW model is effectively the BGK model with 
%\begin{equation}
%\begin{split}
%    \alpha_s(\mu^2)xg(x,\mu^2)\rightarrow &\alpha_s(Q_0^2)xg(x,Q_0^2)\\
%    &=\alpha_s x^{-\lambda}(1-x)^{5.6}.
%\end{split}
%\end{equation}
\commentPending{??}{\color{blue}response: the argument on why we consider $\alpha$ as a constant is removed, as it is not a solid argument.}
%For this reason,
In order to account for the running coupling in Eq.~(\ref{eq:angle-integrated}), 
we assume that $\alpha_s$ in Eq.~(\ref{eq:dipole-gluon}) is
constant for the GBW model, and thus explicitly multiply by the running coupling
%
\begin{equation}
\alpha_s(\mu^2)\fdp(x,k_T^2)\rightarrow \alpha_s(\mu)\frac{\alpha_s\fdp(x,k_T^2)}{0.2},
\end{equation}
where
%
\begin{equation}
\alpha_s(\mu^2)=\frac{1}{\frac{11 C_A-2n_f}{12\pi}\log\left(\frac{\mu^2}{\Lambda_{\mathrm{QCD}}^2 }\right)},
\end{equation}
and  $\Lambda_{\mathrm{QCD}}^2=0.09\;\GeV^2$.

The factor 0.2 is an arbitrary normalization whose effect is absorbed by $\sigma_0$ and hence, bears no importance.
(For a more detailed analysis of the dipole gluon density from the BGK model see Ref.~\cite{Luszczak:2022fkf}\footnote{In Ref.~\cite{Luszczak:2022fkf}, the coupling constant was treated differently, and the large $k_T$-region was matched to the derivative of $xg(x)$, while we have used series transformation to accelerate the high-$k_T$ integration. Their treatment is also different from the original BGK paper~\cite{Bartels:2002cj}. As a consequence their gluon density is positive in the large $k_T$-region and the overall large-$k_T$ behaviour is significantly different.}
)
While there is some ambiguity on what the argument of $\alpha_s(\mu^2)$ should be, we follow Ref.~\cite{Kwiecinski:1997ee} and use
\begin{equation}
	\mu^2=k_T^2+{\kappa'}_t^2+m_f^2+\mu^2_0,
\end{equation}
where we add $\mu^2_0=4\;\GeV^2$ in order to freeze the coupling at low scales.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fits to $F_2$ data}
\label{sec:F2fits}

We fitted the GBW and BGK models to the $F_2$ data from
HERA~\cite{Abt:2017nkc}. A numerical program was written with help of
CERNLIB~(DPSIPG)~\cite{Kolbig:1972zz}, GSL~\cite{GSL}, CUBA~\cite{Hahn:2004fe}
and ROOT~\cite{Brun:1997pa} libraries to evaluate $F_2$. The data were selected
to be in the range $0.045\leq Q^2\leq 650\;\mathrm{GeV^2}$, $x<0.01$.  As with
the previous fit~\cite{Goda:2022wsc}, the $c$ and $b$ flavours were taken into
account with the mass $1.3$ and $4.6\;\mathrm{GeV}$, respectively. As discussed
earlier, we take the light quarks to be massless.  The fitting was performed
using MnMigrad and MnSimplex of ROOT::Minuit2~\cite{James:2004xla}.

We studied the following cases
%
\begin{itemize}
\item GBW model with the fixed coupling in $k_T$-factorization  ($k_T$-GBW),
\item GBW model with the running coupling in $k_T$-factorization (rc-$k_T$-GBW),
\item BGK model in $k_T$-factorization ($k_T$-BGK),
\end{itemize} 
%
and, as a reference, we provide the following results from
Ref.~\cite{Goda:2022wsc};
%
\begin{itemize}
\item GBW model with massless light quarks in the dipole-factorization ($r$-GBW),
\item GBW model with massive light quarks in the dipole-factorization ($r$-GBW-massive),
\item  BGK model in  the dipole-factorization ($r$-BGK).
\end{itemize}

%-----------------------------------------------------------------------------
\begin{table}[t]
\begin{subtable}{\textwidth}
\center\footnotesize
\begin{tabular}{|c|c|c|c|c|}
	\hline - & $\sigma_0 $ [mb] & $x_0 \left(10^{-4}\right)$ & $\lambda$ & $\chi^2/\mathrm{dof}$ \\\hline 
	{\footnotesize $r$-GBW} & 1.907e+01& 2.582e+00& 3.219e-01& 4.438e+00\\\hline 
	{\footnotesize $r$-GBW-massive} & 2.384e+01& 1.117e+00& 3.082e-01& 5.274e+00\\\hline 
	{\footnotesize $k_t$-GBW} & 3.344e+01& 1.333e+00& 3.258e-01& 4.396e+00\\\hline 
	{\footnotesize rc-$k_t$-GBW} & 1.520e+01& 2.648e+00& 3.211e-01& 2.447e+00\\\hline 
\end{tabular}
%\caption{}
\vspace{2mm}
\end{subtable}
\begin{subtable}{\textwidth}
\center\footnotesize
\begin{tabular}{|c|c|c|c|c|c|c|}
	\hline - & $\sigma_0 $ [mb] & $A_g$ & $\lambda_g$ & $C$ & $\mu_0^2 \left[\mathrm{GeV^2}\right]$ & $\chi^2/\mathrm{dof}$ \\\hline 
	{\footnotesize $r$-BGK} & 2.326e+01& 1.181e+00& 8.317e-02& 3.294e-01& 1.873e+00& 1.556e+00\\\hline 
	{\footnotesize $k_t$-BGK} & 3.470e+01& 1.048e+00& 2.205e-01&2.391e-01& 9.954e-01&  1.527e+00\\\hline 
\end{tabular}
%\caption{}
\vspace{2mm}
\end{subtable}
\caption{Fit parameters of respective models. The parameters of the
dipole-factorization cases are from Ref.~\cite{Goda:2022wsc}.}
\label{tab:table}
\end{table}
%-----------------------------------------------------------------------------

The results of the fits are summarized in Tab.~\ref{tab:table}. 
The fit quality of $k_T$-GBW is almost unchanged w.r.t. $r$-GBW, while the rc-$k_T$-GBW shows remarkable improvement, almost halving the $\chi^2$ value.
This is in line with the observation made in Refs.~\cite{Albacete:2004gw,Albacete:2007yr,Albacete:2010sy} that in the BK evolution, the running coupling correction have considerable effects.

Another notable point is that, except for the normalization $\sigma_0$, the parameters are very similar, particularly those of rc-$k_T$-GBW are almost identical to those of $r$-GBW. 
While the GBW model remains almost unaffected, the BGK model seems to show slightly more change. The difference in $\sigma_0$ is similar to that of GBW model, and other parameters changed moderately.

In Fig.~\ref{fig:GBW-BGK}, the plots of the dipole cross section, the dipole
gluon density and the saturation scale for the GBW and BGK models are shown. The
dipole cross section and the gluon density are both normalized by $\sigma_0$ in
order to see the effects  of other parameters better, and the saturation scale
is defined as a ridge of the dipole gluon density in the $(x,k_T^2)$ plane.

In the plots on the left side, one can see the effects of the changes in the parameters of the GBW model. The difference between the rc-$k_T$-GBW and $r$-GBW are negligible, while the $k_T$-GBW is slightly shifted, compared to others, by the change in $x_0$.

On th right side of Fig.~\ref{fig:GBW-BGK}, the same plots for the BGK model are shown.
Unlike in the GBW model, the connection between the differences shown in the plots and the differences in the parameters are not clear. Nevertheless, the difference is more prominent in the small-$x$ region.

Comparisons of the results with the $F_2$ data are shown in Figs.~\ref{fig:GBW-Grid} and~\ref{fig:BGK-Grid}.
In Fig.~\ref{fig:GBW-Grid}, the difference between $r$-GBW and $k_T$-GBW are not visible, while rc-$k_T$-GBW shows sizable difference from the others, particularly in the large-$Q^2$ region. 
Recalling that the parameters of rc-$k_T$-GBW and $r$-GBW are very similar, the difference in $F_2$ is almost entirely due to the coupling constant. The improvement in the fit quality is depicted as a histogram of the $\chi^2$-vale per number of points at the bottom of Fig.~\ref{fig:GBW-Grid}. Here, the improvement in the large-$Q^2$ region is very clear.
In Fig.~\ref{fig:BGK-Grid}, the difference between $r$-BGK and $k_T$-BGK are hardly visible. In the histogram at the bottom, one can see some differences, but they cancel out mostly, making little improvement overall.

\begin{figure}[p]

  \begin{multicols}{2}
  \begin{subfigure}{0.5\textwidth}
    \includegraphics[width=\textwidth]{./plots/GBW-dipole.pdf}
    \includegraphics[width=\textwidth]{./plots/GBW-gluon.pdf}
    \includegraphics[width=\textwidth]{./plots/GBW-saturation.pdf}
    \caption{GBW}
    \label{fig:GBW}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=\textwidth]{./plots/BGK-dipole.pdf}
    \includegraphics[width=\textwidth]{./plots/BGK-gluon.pdf}
    \includegraphics[width=\textwidth]{./plots/BGK-saturation.pdf}
    \caption{BGK}
    \label{fig:BGK}
\end{subfigure}
  \end{multicols}
  \caption{The dipole cross section, the dipole gluon density at $x=10^{-2},\;
  10^{-6}$, and the saturation scale for the GBW (left) and the BGK (right)
  models. Note, the dipole cross section and the gluon density are normalized by
  $\sigma_0^{-1}$.} 
\label{fig:GBW-BGK}
\end{figure}

\begin{figure}[p]
%\begin{subfigure}{\textwidth}
\includegraphics[width=0.49\textwidth]{./plots/Figure_1.png}
\includegraphics[width=0.49\textwidth]{./plots/Figure_2.png}
\includegraphics[width=\textwidth]{./plots/Figure_3.png}
%\end{subfigure}
\caption{Comparison of GBW $F_2$ with HERA data. The histogram shows the $\chi^2$ value per data point in each frame.
 Improvement by the running coupling ($k_T$ run) is clearly visible in the high-$Q^2$region, while the new fit ($k_T$) shows only marginal improvement .  }
\label{fig:GBW-Grid}
\end{figure}

\begin{figure}[p]
%\begin{subfigure}{\textwidth}
\includegraphics[width=0.49\textwidth]{./plots/Figure_2-1.png}
\includegraphics[width=0.49\textwidth]{./plots/Figure_2-2.png}
\includegraphics[width=\textwidth]{./plots/Figure_2-3.png}
%\end{subfigure}
\caption{Comparison of BGK $F_2$ with HERA data. The histogram shows the $\chi^2$ value per data point in each frame. While the overall fit quality remain similar, the fit quality in each frame change noticeably. Particularly the $k_T$-formula case ($k_T$) shows better fit in the small-$Q^2$ region.}
\label{fig:BGK-Grid}
\end{figure}

Let us now take a closer look at the parameter $\sigma_0$. Recall that the
difference between the $k_T$-factorization formula and the dipole-factorization
formula is in $x/z$, Eq.~(\ref{}), and this enters in the GBW formalism as
$\tilde{x}$. It is easy to see that, as $x$ grows, the dipole cross section gets
suppressed. (Keeping in mind the suppression by the photon wave function in the
large-$r$ region.) Such effect was discussed previously in
Ref.~\cite{Kutak:2004ym} in the context of the BK equation. In fact, this
suppression is the motivation given in Ref.~\cite{Golec-Biernat:1998zce} for
such modification of $x$, so that in the small-$Q^2$ limit, the total cross
section remains finite. Since 
%
\begin{equation}
\left(1+\frac{4 m_f^2}{Q^2}\right)\leq\left(1+\frac{k_T^2}{Q^2}+\frac{{\kappa'}_t^2+m_f^2}{\beta(1-\beta)Q^2}\right),
\end{equation}
%
the $k_T$-factorization case receives more suppression. Consequently, the
normalization factor $\sigma_0$ rises to compensate the suppression.  Therefore,
one can understand the change in $\sigma_0$ as a direct consequence of the key
difference between the $k_T$-factorization and the dipole-factorization.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dijet production at EIC}
\label{sec:dijetsEIC}

\begin{figure}[t]
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=\textwidth]{plots/GBWWW1} 
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=\textwidth]{plots/BGKWW1} 
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=\textwidth]{plots/GBWWW2} 
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=\textwidth]{plots/BGKWW2} 
    \end{subfigure}
    \caption{ Weizs\"acker-Williams gluon density at $x=10^{-3}$. Top row: comparison of the dipole-factorization fit and $k_T$-factorization fit results. Bottom row: comparison of the respective models with and without the Sudakov factor, at $\mu=17,\;67 \mathrm{GeV}$.The green dotted line is the KS gluon~\cite{vanHameren:2021sqc}, and the green dashed line is the rcBK gluon~\cite{Albacete:2010sy}. }
    \label{fig:ww}
\end{figure}

The $F_2$ structure function is an inclusive object and is weakly sensitive to
the shape of the gluon density. To probe the latter, we shall now 
apply the dipole cross section obtained to the electron-dijet correlation at the
electron-ion collider~(EIC), following closely the method of
Ref.~\cite{vanHameren:2021sqc}.

We consider dijet production in DIS
%
\begin{equation}
  %\label{eq:}
  e+p\rightarrow e+J_1+J_2+X\,,
\end{equation}
%
At the leading order, in the small-$x$ limit, this process is
dominated by $q\overline{q}$ jets~\cite{Dominguez:2011wm}.  It is
therefore closely related to the dipole picture we discussed earlier. In the Breit frame, where the photon momentum is
given by $q=(0,0,0,Q)$, at the leading order, the jets momentum imbalance
$p_T\equiv\left|p_{1T}+p_{2T}\right|$ equals the gluon transverse momentum
$k_T$, where $p_{1T}$ and $p_{2T}$ are transverse momenta of the jets.  This
makes dijets an interesting process. For the region where $p_T\ll P_T\sim
p_{1T},\,p_{2T}$, one may use power counting to take leading order in $p_T/P_T$,
which leads to the transverse-momentum-dependent~(TMD) factorization.

\commentPending{mention why Sudakov is important for dijet production}\\
{\color{blue} There is a line before Eq.~(\ref{eq:ww-sud}) ?}

In the large-$N_c$ limit of the TMD factorization, there are two types of gluon
densities, namely the dipole gluon density and the Weizs\"acker-Williams (WW)
gluon
density~\cite{Dominguez:2010xd,Dominguez:2011wm,vanHameren:2016ftb,Xiao:2017ggh}.
It was shown in Ref.~\cite{Dominguez:2011wm}, that the dijet process in DIS can
directly probe the WW gluon, $\fww(x,k^2)$, where the differential cross section
factorizes as
%
\begin{equation}
	\frac{d\sigma^{\gamma^*p\rightarrow q\overline{q}X}}{d \mathrm{P.S.}}=\fww(x,k_T^2)H_{\gamma^*g^*\rightarrow q\overline{q}},
        \label{eq:TMD}
\end{equation}
%
with the hard function $H_{\gamma^*g^*\rightarrow q\overline{q}}$ describing
interactions of an off-shell photon with an off-shell gluon producing a
$q\overline{q}$ pair.  $\fww(x,k^2)$, has an interpretation as a number density
of gluons inside a proton, while $\fdp(x,k_T^2)$ does not have such an
interpretation~\cite{Dominguez:2010xd,Dominguez:2011wm,Xiao:2017ggh}.

We carry out our study in the framework of the {\it Improved
Transverse-Momentum-Dependent} (ITMD)
factorization~\cite{Kotko:2015ura,vanHameren:2016ftb}.  The latter is
implemented in the program KaTie~\cite{vanHameren:2016kkz}, which we use to
compute the cross sections.  The ITMD factorization is an generalization of the
TMD factorization where the momentum
imbalance in TMD is restricted to be
small~\cite{Kotko:2015ura,vanHameren:2016ftb}. That is to say ITMD resums
$(Q_s/k_T)^n$ and $(k_T/P_T)^n$~\cite{Kotko:2015ura,vanHameren:2016ftb}, thus
extends the region of applicability up to $k_T\sim P_T$. The difference of
Eq.~(\ref{eq:TMD}) from the regular TMD is that the hard function has an
off-shell gluon $g^*$, thus rendering the $k_T$ dependence in the hard function
as well~\cite{Kotko:2015ura}.

Under the Gaussian approximation and assuming $\theta$-like profile of the
proton, one can
write~\cite{vanHameren:2016ftb,Xiao:2017ggh,Dominguez:2010xd,Dominguez:2011wm}
%
\begin{equation}
\fww(x,k^2)= \frac{C_F}{2\pi^3\alpha_s}\int^\infty_0\frac{dr}{r}J_0(r k) \sdpa(x,r),
\end{equation} 	
with the adjoint dipole cross section
\begin{equation}
\sdpa(x,r)=\sigma_0\left( 1-\left(1-\frac{\sdp(x,r)}{\sigma_0}\right)^{C_A/C_F}\right).
\label{eq:ww}
\end{equation}

For the region where TMD factorization is applicable, $Q_s\sim k_T\ll P_T\sim
Q$, one needs to resum the large Sudakov logarithms $\log(k_T/Q)$, as well as
$\log(1/x)$~\cite{Dominguez:2011wm}. It was shown in
Refs.~\cite{Mueller:2012uf,Mueller:2013wwa,Xiao:2017yya} that consistent
resummation of such logarithms is possible owing to the separation of
corresponding regions.  Resummation of the Sudakov logarithms is achieved by the
formula
%
\begin{equation}
	\fww(x,k^2,\mu^2)= \frac{C_F}{2\pi^3\alpha_s}\int^\infty_0\frac{dr}{r}J_0(r k) e^{-S(r,\mu^2)} \sdpa(x,r),
	\label{eq:ww-sud}
\end{equation}
%
where, we use the Sudakov form factor~\cite{Mueller:2013wwa,Xiao:2017yya},
%
\begin{equation}
	S(r,\mu^2)=\frac{\alpha_s N_c}{4\pi}\ln^2\left(\frac{\mu^2r^2}{4e^{-2\gamma_E}}\right),
\end{equation}
%
in which $\gamma_E$ is the Euler-Mascheroni constant, and we set $\alpha_s=0.2$. 
Following Ref.~\cite{vanHameren:2021sqc}, we study the azimuthal correlations of
jets and the final state electron in DIS , where
it was argued that this observable is sensitive to the soft emissions and the
saturation effects. In this study we focus only on the proton case.  The
kinematical cuts suggested in Ref.~\cite{vanHameren:2021sqc} were
%
\begin{align*}
	E_e&=15\GeV& E_p&=135\GeV& Q^2&>1\GeV^2\\
	0.1<&\nu<0.85&\Delta R_{\mathrm{Breit}}&<1&p^{\mathrm{Breit}}_{1,2\,T}&>3\GeV\\
	&&-4<&y_{1,2\,\mathrm{lab}}<-1.&&
\end{align*}

\comment{Question: normalization}

Grids of the Weizs\"acker-Williams gluon density were produced by evaluating
Eqs.~(\ref{eq:ww}) and~(\ref{eq:ww-sud}).  The gluon density at $x=10^{-3}$ is
plotted in Fig.~\ref{fig:ww} with hard-scale-independent KS
gluon~\cite{vanHameren:2021sqc,Abdulov:2021ivr} and running-coupling BK~(rcBK)
gluon density~\cite{Balitsky:2006wa,Albacete:2010sy}. Clearly, the GBW and BGK models falls much more quickly
than the KS and rcBK gluon density. In general, expected behaviour in the
large-$k_T$ region is $\sim k_T^{-2}$~\cite{Dominguez:2010xd,Dominguez:2011wm},
while $\sigma_{\mathrm{GBW}}$ behaves like $\sim e^{-k_T^2}$.  As in
Ref.~\cite{vanHameren:2021sqc}, the Sudakov factor enhances in the small-$k_T$
region and suppresses in the large-$k_T$ region. In other words, it broadens the
distribution. In comparison to the result of Ref.~\cite{vanHameren:2021sqc}, the
effect of broadening by the Sudakov factor is significantly more pronounced in
the case of the GBW and BGK models. The hard-scale-dependent GBW and BGK models,
as a consequence, become closer to the KS and rcBK gluons, \textit{cf.} Fig.~2
of Ref.~\cite{vanHameren:2021sqc}.

Figs.~\ref{fig:je-breit} and~\ref{fig:je-lab} show electron-jets azimuthal
correlation in the Breit and in the lab frame respectively.  In the top row of Fig.~\ref{fig:je-breit}, we see, for both the
GBW and BGK models, better agreements of results with the KS and rcBK for the
new $k_T$-factorization fits. However the overall normalization of the gluon
density depends on the coupling $\alpha_s$, which we assumed to be 0.2.
Nevertheless, it shows clearly the effect of the parameter $\sigma_0$. In the
middle and the bottom row, it shows the effects of the Sudakov form factor,
which qualitatively agrees with that of Ref.~\cite{vanHameren:2021sqc}, by
lowering the cross section.

Fig.~\ref{fig:je-lab} shows the electron-jets
correlation in the lab frame. Here, the difference between KS and GBW and BGK
is more prominent, while rcBK shows similar pattern to the GBW and BGK models.
The effects of the Sudakov factor is similar to those
in Ref.~\cite{vanHameren:2021sqc} at relatively high $\Delta \phi$, while at
smaller $\Delta \phi$ region the effect is reversed ({\it i.e.} the cross
sections were slightly lowered in Ref.~\cite{vanHameren:2021sqc}, while here,
they are significantly increased).

\begin{figure}[p]
	\begin{subfigure}{0.5\textwidth}
	\includegraphics[width=\textwidth]{plots/plotGBW1} 
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\includegraphics[width=\textwidth]{plots/plotBGK1} 
	\end{subfigure}
%	\caption{ }
%\end{figure}
%\begin{figure}[t]
	\begin{subfigure}{0.5\textwidth}
	\includegraphics[width=\textwidth]{plots/plotGBW2}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\includegraphics[width=\textwidth]{plots/plotGBW3}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\includegraphics[width=\textwidth]{plots/plotBGK2}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\includegraphics[width=\textwidth]{plots/plotBGK3}
	\end{subfigure}
	\caption{ Azimuthal correlation of the jets and the scattered electron in the Breit frame. Top: Comparison of dipole-factorization fit and $k_T$-factorization fit. Middle \& Bottom: Effect of the Sudakov form factor. The green dotted line is the KS gluon~\cite{vanHameren:2021sqc}, and the green dashed line is the rcBK gluon~\cite{Albacete:2010sy}.}
	\label{fig:je-breit}
\end{figure}

\begin{figure}[p]
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[width=\textwidth]{plots/plotGBW1Lab} 
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[width=\textwidth]{plots/plotBGK1Lab} 
	\end{subfigure}

	\begin{subfigure}{0.5\textwidth}
		\includegraphics[width=\textwidth]{plots/plotGBW2Lab}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[width=\textwidth]{plots/plotGBW3Lab}
	\end{subfigure}

	\begin{subfigure}{0.5\textwidth}
		\includegraphics[width=\textwidth]{plots/plotBGK2Lab}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[width=\textwidth]{plots/plotBGK3Lab}
	\end{subfigure}
\caption{ Azimuthal correlation of the jets and the scattered electron in the lab frame. Top: Comparison of dipole-factorization fit and $k_T$-factorization fit. Middle \& Bottom: Effect of the Sudakov form factor. The green dotted line is the KS gluon~\cite{vanHameren:2021sqc}, and the green dashed line is the rcBK gluon~\cite{Albacete:2010sy}.}
\label{fig:je-lab}
\end{figure}

Finally, Fig.~\ref{fig:jj-breit} shows the jet-jet correlations in the Breit
frame. Again, the GBW and BGK models exhibit considerable deviation from the KS
gluon. The difference from the previous plot is the disagreement of the GBW/BGK
models and the rcBK in the small-$\Delta\phi$ region. Similarly to the previous
plots, the Sudakov factor affects the models somewhat differently from the KS
gluon in Ref.~\cite{vanHameren:2021sqc}.  The effect enhances the cross section
considerably in the small-$\Delta\phi$ region, making it closer to KS gluon
result.

The results shown in Figs.~\ref{fig:je-lab} and~\ref{fig:jj-breit} are
natural as back-to-back configuration in the respective observable corresponds
to the small-$k_T$ region of gluon densities, and as it can be seen clearly in
Fig.~\ref{fig:ww}, the GBW and BGK gluons do not fare well in the large-$k_T$
region.  That is to say that the enhancement in the small-$\Delta\phi$ region is
a direct consequence of the broadening by the Sudakov factor.

\begin{figure}[p]
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[width=\textwidth]{plots/plotGBW1Jets} 
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[width=\textwidth]{plots/plotBGK1Jets} 
	\end{subfigure}

	\begin{subfigure}{0.5\textwidth}
		\includegraphics[width=\textwidth]{plots/plotGBW2Jets}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[width=\textwidth]{plots/plotGBW3Jets}
	\end{subfigure}

	\begin{subfigure}{0.5\textwidth}
		\includegraphics[width=\textwidth]{plots/plotBGK2Jets}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[width=\textwidth]{plots/plotBGK3Jets}
	\end{subfigure}
\caption{Azimuthal correlation of the jets in the Breit frame. Top: Comparison of dipole-factorization fit and $k_T$-factorization fit. Middle \& Bottom: Effect of the Sudakov form factor. The green dotted line is the KS gluon~\cite{vanHameren:2021sqc}, and the green dashed line is the rcBK gluon~\cite{Albacete:2010sy}. }
\label{fig:jj-breit}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary}

We have fitted the GBW and BGK saturation models to HERA data~\cite{Abt:2017nkc}
using the $k_T$-factorization for the structure function $F_2$.  The main
difference between the $k_T$- and the dipole-factorizations is an argument of
the gluon, that is $x/z$ appearing in the former being replaced by
$x(1+4m_f^2/Q^2)$ in the latter. 
%
In fact the massive light quarks used in
Ref.~\cite{Golec-Biernat:1998zce} partially simulate the factor $1/z$, and our 
fit result indicates such effect, as expected.

It is quite remarkable that with such simplicity, the dipole-factorization can
reproduce the result of the $k_T$-factorization formula well. We have found that
the only major difference is in the normalization parameter $\sigma_0$ which
increases significantly for the $k_T$-factorization case. We have also observed
that the explicit inclusion of the running coupling has significant effect on
the fit quality, particularly in the large-$Q^2$ region, where the GBW model
performs poorly.

We have applied the new result from our fits for the prediction of the dijet
process in DIS at EIC. Additionally, effects of the Sudakov form factor were
investigated for that process. Results of the electron-jets correlation in the
Breit frame agree qualitatively to those of Ref.~\cite{vanHameren:2021sqc}.
Other results, namely the electron-jets in the lab frame and the jet-jet
correlation in the Breit frame, show considerable effects of the Sudakov form
factor which broaden the gluon density.

\comment{A few more closing sentences. (SS)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgment}
We are grateful to Krzysztof Golec-Biernat for numerous useful discussions and
for careful reading of the manuscript, and Piotr Kotko and Andreas van Hameren
for the clarification of details of Ref.~\cite{vanHameren:2021sqc} and 
usage of the KaTie package. The project is partially supported by the European
Union’s Horizon 2020 research and innovation program under grant agreement No.
824093.  TG and SS are partially supported by the Polish National Science Centre
grant no. 2017/27/B/ST2/02004.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{$I_i$ in Eq.~(\ref{eq:angle-integrated})}

The functions $I_1,\, I_2,\,  I_3,\,  I_4$ used in
Eq.~(\ref{eq:angle-integrated}) are defined  as~\cite{Kimber:2001uaa}
%
\begin{equation}
	\begin{split}
		\frac{I_1}{2\pi}=\frac{N_1N_2+N_3^2}{\left( N^2_1+2N_1N_2+N_3^2\right)^{3/2}}, &\hspace{1cm}
		\frac{I_2}{2\pi}=\frac{N_3-(1-2\beta)N_1}{(N_1+N_4)\sqrt{ N^2_1+2N_1N_2+N_3^2}},\\
		\frac{I_3}{2\pi}=\frac{N_1+N_2}{\left( N^2_1+2N_1N_2+N_3^2\right)^{3/2}},&\hspace{1cm}
		\frac{I_2}{2\pi}=\frac{(1-\beta)}{(N_1+N_4)\sqrt{ N^2_1+2N_1N_2+N_3^2}},
	\end{split}
\end{equation}
%
for
%
\begin{equation}
	\begin{split}
		N_1\equiv\beta(1-\beta)Q^2+m_f^2, &\hspace{2cm}
		N_2\equiv{\kappa'}_t^2+(1-\beta)^2k_T^2,\\
		N_3\equiv{\kappa'}_t^2-(1-\beta)^2k_T^2, &\hspace{2cm}
		N_4\equiv{\kappa'}_t^2+\beta(1-\beta)k_T^2.
	\end{split}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\clearpage
%\printbibliography

\bibliographystyle{unsrt}
%\bibliographystyle{apalike}
\bibliography{draft}

%\begin{thebibliography}{10}
%\bibliography{draft}
%\bibliographystyle{unsrt}

%\input{./draft-references}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% !!! DO NOT DELETE !!!
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\cite{GSL}
%\bibitem{GSL}
%M.~Galassi, J.~Davies, J.~Theiler, B.~Gough, G.~Jungman,
%"GNU Scientific Library - Reference Manual, Third Edition, for GSL Version 1.12 (3. ed.)",
%(2009),
%isbn:~978-0-9546120-7-8
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\end{thebibliography}
\end{document}